A new method to estimate dark matter halo concentrations

### 2015-02-23 paper/article.tex
new references added to the intro

\newcommand{\apj}{ApJ}
\newcommand{\apjs}{ApJS}
\newcommand{\apjl}{ApJL}
\newcommand{\aj}{AJ}
\newcommand{\mnras}{MNRAS}
\newcommand{\mnrassub}{MNRAS accepted}
\newcommand{\aap}{A\&A}
\newcommand{\aaps}{A\&AS}
\newcommand{\araa}{ARA\&A}
\newcommand{\nat}{Nature}
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}
\newcommand{\pasj}{PASJ}
\newcommand{\avg}[1]{\langle{#1}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$~}\fi}}
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}
\newcommand{\hs}{{\hspace{1mm}}}
\newcommand{\muavg}{\vert\langle\cos\theta\rangle\vert}
% definition to produce a "less than or similar to" symbol
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}

@@

\section{Introduction}
\label{sec:introduction}
In the concordance cosmology paradigm the the matter content of the
Universe is dominated by dark matter which behaves as a collisionless
fluid under the influence of gravity. In the last three decades
numerical experiments have made possible the simulation of dark matter
dominated universes, yielding valuable insights on the large scale
structure formation process as depicted in observations.

One of the most striking results of these simulations is that dark
matter clumps on galactic scales follow a universal density profile
which in a first approximation is spherical symmetric and only
dependent on the radial coordinate. These profiles seem to be
universal; independent of the cosmological parameters and self-similar
for different spatial scales after an adequate re-scaling is
app plied.

One the most common parameterization of this density is known as the
Navarro-Frenk-White (NFW) profile \citep{NFW}. This profile is a
double power law in radius, where the transition between the two
happens at the so-called scale radius $r_s$. The ratio between the
scale radius and the virial radius of the halo $R_v$ is known as the
concentration $c=R_V/r_s$, a quantity that has been found to be
correlated with the halo mass.

The relationship between halo mass and concentration is in principle
accessible to observations and provides a potential test of LCDM on
galactic scales. With this promise a great deal of effort has been
invested in calibrating this relationship with simulations but also
finding the best possible way to constraint it with observations.

>>>>

\newcommand{\apj}{ApJ}
\newcommand{\apjs}{ApJS}
\newcommand{\apjl}{ApJL}
\newcommand{\aj}{AJ}
\newcommand{\mnras}{MNRAS}
\newcommand{\jcap}{JCAP}
\newcommand{\mnrassub}{MNRAS accepted}
\newcommand{\aap}{A\&A}
\newcommand{\aaps}{A\&AS}
\newcommand{\araa}{ARA\&A}
\newcommand{\nat}{Nature}
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}
\newcommand{\pasj}{PASJ}
\newcommand{\avg}[1]{\langle{#1}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$~}\fi}}
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}
\newcommand{\hs}{{\hspace{1mm}}}
\newcommand{\muavg}{\vert\langle\cos\theta\rangle\vert}
% definition to produce a "less than or similar to" symbol
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}

@@

\section{Introduction}
\label{sec:introduction}
In the concordance cosmology paradigm the the matter content of the
Universe is dominated by dark matter which behaves as a collisionless
fluid under the influence of gravity. In the last three decades
numerical experiments have made possible the simulation of dark matter
dominated universes, yielding valuable insights on the large scale
structure formation process as depicted in observations
\citep{Springel2005,2011ApJ...740..102K}.

One of the most striking results of these simulations is that dark
matter clumps on galactic scales follow a universal density profile
which in a first approximation is spherical symmetric and only
dependent on the radial coordinate. These profiles seem to be to a
very good approximation universal \citep{NFW,Taylor2001}; independent
of the cosmological parameters and self-similar for different spatial
scales after an adequate re-scaling is applied.

One the most common parameterization of this density is known as the
Navarro-Frenk-White (NFW) profile \citep{NFW}. This profile is a
double power law in radius, where the transition between the two
happens at the so-called scale radius $r_s$. The ratio between the
scale radius and the virial radius of the halo $R_v$ is known as the
concentration $c=R_v/r_s$, a quantity that has been found to be
correlated with the halo mass.

In spite of the fact that high resolution simulations of
Milky Way sized dark matter halos \citep{Navarro2010} show that the
universality property is not  perfect and that a better fitting
parameterization to the radial density profile is provided by the
Einasto profile \citep{Einasto1965}, the NFW and its concentration
have become a standard and useful way to describe the structure of
dark matter halos.

The relationship between halo mass and concentration is in principle
accessible to observations and provides a potential test of LCDM on
galactic scales. 
With this promise a great deal of effort has been
invested in calibrating this relationship with simulations
\citep{Neto2007,Duffy2008,Munoz2011,Prada2012,Ludlow2014} but also
finding the best possible way to constraint it with observations
\citep{Buote2007,Comerford2007,Mandelbaum2008,Giocoli2014,Foex2014,Shan2015}.
